Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-04 robdml.tex
tex

Given a set of data points, a \emph{metric} gives a distance value between each
pair of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
As an extensively studied task~\cite{revisiting}, DML has a wide range of
applications such as image retrieval~\cite{imagesim2} and face
recognition~\cite{facenet,domainface}, as well as a wide influence on the core
idea of other areas such as self-supervised learning~\cite{dmlreality}.

Despite the advancements in this field thanks to deep neural
networks, recent studies find DML models vulnerable to
adversarial attacks, where an imperceptible perturbation can incur unexpected
retrieval, or covertly manipulate the results~\cite{advrank,advorder}.
%
Such vulnerability raises security, safety, and fairness concerns in the
DML applications.
%
For example, impersonation or recognition evasion are possible for a
vulnerable DML-based face-identification system.
%
To counter the attacks (\ie, mitigating the vulnerability), the
\emph{adversarial robustness} of DML models has to be improved via defense.

>>>>

Given a set of data points, a \emph{metric} gives a distance value between each
pair of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
As an extensively studied task~\cite{revisiting,dmlreality}, DML has a wide
range of applications such as image retrieval~\cite{imagesim2} and face
recognition~\cite{facenet,domainface}, and widely influences some other areas
such as self-supervised learning~\cite{dmlreality}.

Despite the advancements in this field thanks to deep learning, recent studies
find DML models vulnerable to adversarial attack, where an imperceptible
perturbation can incur unexpected retrieval result, or covertly change the
rankings~\cite{advrank,advorder}.
%
Such vulnerability raises security, safety, and fairness concerns in DML
applications.
%
For example, impersonation or recognition evasion are possible on a vulnerable
DML-based face-identification system.
%
To counter the attacks (\ie, mitigating the vulnerability), the
\emph{adversarial robustness} of DML models has to be improved via defense.
